---
title: "ggplot"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggridges)

```

```{r load_data, cache = TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

```
##create a ggplot
```{r}
ggplot(weather_df, aes(x= tmin, y = tmax)) + #definding the dataset itself, what's the x-axis, and what's the y-axis;
  geom_point()
```
saving initial plots

```{r}
#weather_df %>% filter(name ==)
scatterplot = 
   weather_df %>% 
  ggplot(aes(x= tmin, y = tmax)) + geom_point()

scatterplot
```

adding color

```{r}
weather_df %>% 
  ggplot(aes(x= tmin, y = tmax)) + 
  geom_point(aes(color=name), alpha = 0.4) #alpha statemtn: adjusting the transparency, range[0,1]

```
why do 'aes' positions matter?

```{r}
weather_df %>% 
  ggplot(aes(x= tmin, y = tmax)) + 
  geom_point(aes(color=name), alpha = 0.4) +
  geom_smooth(se.=FALSE) #giving a smooth curve about the scatterplot; #se.=FALSE about confidence interval
```
vs. 

```{r}
weather_df %>% 
  ggplot(aes(x= tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.4) +
  geom_smooth(se = FALSE)  #giving a smooth curve about the scatterplot; #se =FALSE:get rid of the confidence interval shown in the graph
  facet_grid(~ name) #facet_grid: separate the data by name
```
```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE)
```
```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_smooth(size =2, se = FALSE) 
```

2d density

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_bin2d()
  #geom_hex(): after install.packages("hexbin")

```
##more kinds of plots!
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) + #color: tells you the bar outside, if you want to color the inside part, use fill
  geom_histogram() +
  facet_grid(~name)

```
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_density(alpha = 0.3) 
  #facet_grid(~name)
```
```{r}
  ggplot(weather_df, aes(x = name, y = tmax)) + geom_boxplot()
```
```{r}
ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  stat_summary(fun.y = median, geom = "point", color = "blue", size = 4)
```

```{r}
weather_df %>% 
ggplot(aes(x = tmax, y = name)) + 
  geom_density_ridges(scale = .85)
```

## saving a plot

```{r}
ggp_ridge_temp = 
  weather_df %>% 
  ggplot(aes( x= tmax, y= name)) +
  geom_density_ridges()

ggsave("ggplot_temp_ridge.pdf",
       ggp_ridge_temp)
```

```{r, fig.width = 12}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(alpha = 0.4) +
  geom_smooth(se = FALSE)


```

## Making more plots
```{r}
weather_df %>% 
  ggplot (aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5)
```

adding labels:

```{r}
weather_df %>% 
  ggplot (aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature Plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from NOAA via rnoaa package"
    )
```

x axis tick marks etc

```{r}
weather_df %>% 
  ggplot (aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature Plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from NOAA via rnoaa package"
  ) +
  scale_x_continuous(
    breaks = c(-15, -5, 20),
    labels = c("-15C", "-5(still too cold)", "20")
  ) +
  scale_y_sqrt()
#scale_y_continuous( trans = "sqrt") doing the same thing: transforming the function into the sqaure root

```
change the default color using scale_color_hue
```{r}
weather_df %>% 
  ggplot (aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature Plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from NOAA via rnoaa package"
  ) +
  scale_color_hue(
    name = "Weather Station", #naming the color reference table, called legand
    h = c(50,250) #changing the value here to adjust the color of the plot
  )
```
```{r}
ggp_base = 
weather_df %>% 
  ggplot (aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature Plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from NOAA via rnoaa package"
  ) +
 viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
 )

ggp_base
```

##themes

```{r}
ggp_base +
  theme (legend.position = "bottom") #put the legend on the bottom of the plot
```
```{r}
ggp_base +
  theme_minimal() + #erasing the grey background color
  theme(legend.position = "bottom")
#the order matters! if do the theme(legend.position) first, the plot will first put the legend on the bottom, but when it goes to the next argument, the theme_Minimal, the legend will go back on the right side of the plot.
```


```{r}
weather_df %>% 
  filter(is.na(tmin))
```

## more than one dataset
```{r}
central_park = 
  weather_df %>% 
  filter (name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter (name == "Waikiki_HA")

ggplot(data = waikiki, aes (x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp)) +
  geom_line(data = central_park)
```

bried aside about colors

```{r}
waikiki %>% 
  ggplot(aes (x = date, y = tmax)) + #to make the color red
  geom_point(alpha = 0.5, color = "red") 
  # to get the specific color plots by stating it on geom_point under color statement

```

multiple-panel plots
```{r}
library(patchwork)

ggp_scatter = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_density = 
  weather_df %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_density() #geom_density: creating the density curve

ggp_box = 
  weather_df %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot() #geom_boxplot: creating the boxplot



ggp_scatter +ggp_density
```

Data Manipulation
```{r}
weather_df %>% 
  mutate(
    name = factor(name),
    name = fct_relevel(name, "Waterhole_WA", "Waikiki_HA", "CentralPark_NY")
  ) %>% 
  ggplot(aes(x = name, y = tmax, color = name)) +
  geom_boxplot()
```

reorder instead of relevel
```{r}
weather_df %>% 
  mutate(
    name = factor(name),
    name = fct_reorder(name, tmax)
  ) %>% 
  ggplot(aes(x = name, y = tmax, color = name)) +
  geom_boxplot()
```
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name))+ 
  geom_point()
```
## reconstructure then plot

```{r}
weather_df %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "observation",
    values_to = "temperature"
  ) %>% 
  ggplot (aes(x = temperature, fill = observation)) +
  geom_density(alpha = 0.5) + #change the transparency
  facet_grid(~name) + #straty the plots by names 
  theme(legend.position = "bottom") #put the legend on the bottom
```
```{r}
pup_data = 
  read_csv("./data_import_examples/FAS_pups.csv", col_types = "ciiiii") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litter_data = 
  read_csv("./data_import_examples/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3) %>%
  mutate(wt_gain = gd18_weight - gd0_weight,
         day_of_tx = as.numeric(day_of_tx))

fas_data = left_join(pup_data, litter_data, by = "litter_number") 

fas_data %>% 
  select(sex, dose, day_of_tx, pd_ears:pd_walk) %>% 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome", 
    values_to = "pn_day") %>% 
  drop_na() %>% 
  mutate(outcome = forcats::fct_reorder(outcome, day_of_tx, median)) %>% 
  ggplot(aes(x = dose, y = pn_day)) + 
  geom_violin() + 
  facet_grid(day_of_tx ~ outcome)
```
```{r}
fas_data %>% 
  drop_na() %>% 
  ggplot(aes(x = dose, y = pd_ears)) +
  geom_violin() +
  facet_grid(day_of_tx ~ .)
```

